Reproducible Research in R: A tutorial on how to do the same thing more than once

### 2021-10-07 manuscript.Rmd
grammerly on introduction

Scientists increasingly strive to make research data, materials, and analysis code openly available.
Sharing these types of digital research products has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
To some extent these potential benefits are contingent on whether these digital research products are reproducible. 
Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
The credibility of empirical research results hinges on their objectivity.
Objectivity in this context means "in principle it [the research finding] can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper].
Only a reproducible result can be tested and understood by anyone.
Therefore, reproducibility has long been considered integral to empirical research.
Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges (e.g., missing software dependencies or ambiguous documentation of the exact computational steps taken), and there is a dire need for reproducibility standards.
With the technological advancement, it is now possible to make all digital products related to a research project available in a way that enables automatic reproduction of the research project with minimal effort.

With this paper we pursue two aims.
First, we want to introduce researchers to a notion of automated reproducibility that requires, apart from the setup of software, no manual steps. 
Second, we discuss opportunities such automated reproducibility affords to improve not only how research outputs are created and recreated, but also how research itself is conducted.
With regard to the first goal, we discuss how to address four common threats to reproducibility with tools originating from software engineering [see @Peikert2019 for details].
Then we present a tutorial on how to employ these tools to achieve automated reproducibility.
A single tutorial cannot comprehensively introduce the reader to the individual tools, but this tutorial is intended to help readers to build a coherent workflow.
This tutorial is aimed at researchers who regularly write code to analyze their data and who are willing to make relevant code, data, and materials available, either publicly or on request.
Ideally, our reader already has created a dynamic document at some point in time (e.g. with RMarkdown or Jupyter) and attempted to track a file with version control (e.g. with Git). 
The R package `repro` supports researchers in setting up the required software and in adopting this workflow.
We present automated reproducibility as best practice, a goal that is not always archived because of limited resources, technical restrictions or practical considerations but is worth striving for.
If, due to whatever reason, some parts of the workflow proof to be unfeasible, we encourage the reader to find their own suitable solutions and to concentrate on the feasible parts.

In pursuit of the second aim, we present a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon an implementation of this reproducible workflow, the so-called *preregistration as code* (PAC).
PAC involves preregistering the intended analysis code and the major part of the final scientific report as a dynamic document, including typical sections like Introduction, Methods, and Results.
The resulting dynamic document closely resembles the final manuscript, but temporarily uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
With these placeholders, authors preregister the exact presentation of their results without having seen real data yet.
Once the empirical data are available, they replace the simulated data, the results are then updated automatically, and the discussion can be written to finalize the report.

@@

Scientific organizations and funding bodies increasingly demand transparent sharing of digital research products, and researchers are increasingly willing to do so.
However, although the sharing of such digital research products is a necessary condition for reproducibility, it is not a sufficient one.
This was illustrated by an attempt to reproduce results from open materials in the journal *Cognition* [@hardwicke2018]. 
Out of 35 published articles with open code and data, results of 22 articles could be reproduced but in 11 of these cases, further assistance from the original authors was required.
For 13 articles, at least one outcome could not be reproduced—--even with the original authors' assistance.
Another study of 62 Registered Reports found that only 41 had data available, and 37 had analysis scripts available [@obels2020].
The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
These failed attempts to reproduce highlight the need for widely accepted reproducibility standards, because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
If digital research products are available but not reproducible, their added value is limited.

@@

1.	*Multiple inconsistent versions of code, data, or both*; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code could have been modified during writing of a paper because a bug was removed at some point in time.
It may then be unclear which version of code and data was used to produce some reported set of results.
2.	*Copy-and-paste errors*; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result.
3.	*Undocumented or ambiguous order of computation*; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another).
4.	*Ambiguous software dependencies*; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.

@@

*Version control* prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the lifetime (and in particular, in the development phase) of a research project.
Version control allows a clear link between which results were generated by which version of code and data.
This is important for reproducibility because only if it is obvious which version of data and code produced which results they can be said to be reproducible.
We recommend using *Git* for version control, because of its widespread adoption in the R community.

@@

Dynamic document generation integrates both steps.
Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allowing interested readers to gain a deeper understanding of the contents [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
*R Markdown* uses `Markdown` for formatting text and R (or other programming languages) for writing the statistical analysis.
`Markdown` is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
This way, `Markdown` does not need any specialized software for editing.
It is highly fool-proof [unlike, for example, `LaTeX` @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
We suggest using `Markdown` for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas, to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
`R Markdown` extends regular `Markdown` by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a `Markdown` document.
Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
This allows the creation of (conditionally) formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
@xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of `R Markdown`.

One would hope that sharing a dynamic document would allow individuals (including one's future self) to simply download and reproduce a given analysis.
While version control and dynamic document generation are becoming more common, we have argued that two more components are required and that each component alone is unlikely to guarantee reproducibility [@Peikert2019; @vanlissa2020worcs].
In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.

To automatically resolve dependencies between project files, we rely on *dependency tracking*.
In essence, researchers provide a collection of *computational recipes*.
A computational recipe describes how inputs are processed to deterministically create a specific output in a way that is automatically executable.
The concept of computational recipes is central to our understanding of reproducibility because it enables a unified way to automatically reproduce a project.
Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).
One noteworthy property of computational recipes is that they require a textual representation and the ability of a program to make sense of this textual representation.
A workflow that includes, e.g. renaming variable names by hand in a graphical spreadsheet application is therefore incompatible with automated reproducibility.
Another property of a computational recipe is that the same inputs should always result in the same outputs.
Fo most computer code (given the same software is used) this property is fulfilled, however, one noteworthy exception are pseudo random numbers.
Whenever random numbers are used in a computation, this computation is only reproducible when the same random numbers are used.
To ensure identical random numbers, users may fix the state of the random number generate with a so called seed (e.g. in R with `set.seed()`) but they also need to guarantee that the pseudo random number generator is unchanged [see @Peikert2019].

@@

There are two targets, the final rendered report (`manuscript.pdf`, l.1) and the simulation results (`simulation_results.csv`, l.4).
Each target is followed by a colon and a list of requirements.
If a requirement is newer than the target, the recipe will be executed to rebuild the target.
If a requirement does not exist, `Make` uses a recipe to build the requirement first before building the target.
Here, if one were to build the final `manuscript.pdf` by rendering the `R Markdown` with the command shown in l.2, `Make` would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation first before rendering the manuscript.
This ensures that the simulated data are present before the manuscript is built, and that the simulation is rerun and the manuscript is rebuilt if the simulation code was changed.
`Make` therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] does not contain this element of dependency tracking explicitly but its strict structure of only containing one definite `R Markdown` still makes dependencies between files unambiguous.

@@

A version controlled dynamic document with dependency tracking may still rely on external software.
Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide all dependent software packages and system libraries via *software management*.
One comprehensive approach to software management is containerization.
The central idea is that by
"packaging the key elements of the computational  environment  needed  to  run  the  desired  software the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p. 174].

@@

Note that the software environment of the Docker image is completely separate from software installed on your computer.
This separation is excellent for reproducibility, but takes some getting used to.
For example, it is important to realize that software available on your local computer will *not* be accessible within the confines of the `Docker` image.
Each dependency that you want to use within the `Docker` image must be explicitly added as a dependency.
Furthermore, using `Docker` may require you to install software on an operating system that may not be familiar to you.
The images supplied by @boettigerIntroductionRockerDocker2017, for example, are based on Linux.

There are two ways to build a `Docker` image.
First users can manually install whatever software they like from within the virtual environment.
Such manually build environment can still be ported to all computers which support `Docker`.
We, however, prefer the second way of building images automatically from a textual description called `Dockerfile`.
Because the `Dockerfile` clearly describes which software is installed how the installation process can be repeated automatically.
Users can therefore quickly change the software environment, e.g. update to another R version or given package version.

Packaging all required software in such an image requires considerable amounts of storage space.
Two major strategies help to keep the storage requirements reasonable.
One strategy is to rely on pre-made images, which are maintained by a community for particular purposes.
For example, there are pre-made images that only include what is necessary for R, based on Ubuntu containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
The image for this project includes Ubuntu, R, RStudio, LaTeX as well as a variety of R packages like tidyverse [@tidyverse] and all its dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
`Dockerfiles` are tiny (the [`Dockerfile`](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
For proper archival, we therefore recommend to store a complete image of the software environment, in addition to the `Dockerfile`.
A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] relies for software management on the R package `renv`[@R-renv] which is more lightweight and easier to use than `Docker` on the one hand, but not as comprehensive on the other hand because it only snapshots the R packages instead of all software used.

To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using `R Markdown`), version control (using `Git`), dependency tracking (using `Make`), and Software management (using `Docker`).
While `R Markdown` and `Git` are well integrated into the R environment through RStudio, `Make` and `Docker` require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable obstacle to the acceptance and implementation of the workflow.
To remove this obstacle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
These projects will remain reproducible irrespective of whether `repro` remains accessible in future.
Users do not have to have `repro` installed to reproduce a project;
in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.

>>>>

Scientists increasingly strive to make research data, materials, and analysis code openly available.
Sharing these digital research products can increase scientific efficiency by enabling researchers to learn from each other, reuse materials, and increase scientific reliability by facilitating the review and replication of published research results.
To some extent, these potential benefits are contingent on whether these digital research products are reproducible. 
Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
The credibility of empirical research results hinges on their objectivity.
Objectivity in this context means "in principle it [the research finding] can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper].
Only a reproducible result can be tested and understood by anyone.
Therefore, reproducibility has long been considered integral to empirical research.
Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
This is because making a research project with all code, data, and materials reproducible comes with various challenges (e.g., missing software dependencies or ambiguous documentation of the exact computational steps taken), and there is a dire need for reproducibility standards.
With technological advancement, it is now possible to make all digital products related to a research project available in a manner that enables automatic reproduction of the research project with minimal effort.

With this paper, we pursue two aims.
First, we want to introduce researchers to a notion of automated reproducibility that requires, apart from the setup of the software, no manual steps. 
Second, we discuss opportunities such automated reproducibility affords to improve not only how research outputs are created and recreated but also how research itself is conducted.
With regard to the first goal, we discuss how to address four common threats to reproducibility with tools originating from software engineering [see @Peikert2019 for details].
Then we present a tutorial on how to employ these tools to achieve automated reproducibility.
A single tutorial cannot comprehensively introduce the reader to the individual tools, but this tutorial is intended to help readers to build a coherent workflow.
This tutorial is aimed at researchers who regularly write code to analyze their data and who are willing to make relevant code, data, and materials available, either publicly or on request.
Ideally, our reader already has created a dynamic document at some point in time (e.g. with RMarkdown or Jupyter) and attempted to track a file with version control (e.g. with Git). 
The R package `repro` supports researchers in setting up the required software and in adopting this workflow.
We present automated reproducibility as best practice, a goal that is not always archived because of limited resources, technical restrictions or practical considerations but is worth striving for.
If, due to whatever reason, some parts of the workflow prove to be unfeasible, we encourage the reader to find their own suitable solutions and to concentrate on the feasible parts.

In pursuit of the second aim, we present a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon implementing this reproducible workflow, the so-called *preregistration as code* (PAC).
PAC involves preregistering the intended analysis code and the major part of the final scientific report as a dynamic document, including typical sections like Introduction, Methods, and Results.
The resulting dynamic document closely resembles the final manuscript but temporarily uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
With these placeholders, authors preregister the exact presentation of their results without having seen real data yet.
Once the empirical data are available, they replace the simulated data, the results are then updated automatically, and the discussion can be written to finalize the report.

@@

Scientific organizations and funding bodies increasingly demand transparent sharing of digital research products, and researchers are increasingly willing to do so.
However, although the sharing of such digital research products is a necessary condition for reproducibility, it is not a sufficient one.
This was illustrated by an attempt to reproduce results from open materials in the journal *Cognition* [@hardwicke2018]. 
Out of 35 published articles with open code and data, results of 22 articles could be reproduced, but in 11 of these cases, further assistance from the original authors was required.
For 13 articles, at least one outcome could not be reproduced—--even with the original authors' assistance.
Another study of 62 Registered Reports found that only 41 had data available, and 37 had analysis scripts available [@obels2020].
The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
These failed attempts to reproduce highlight the need for widely accepted reproducibility standards because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
If digital research products are available but not reproducible, their added value is limited.

@@

1.	*Multiple inconsistent versions of code, data, or both*; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code could have been modified during the writing of a paper because a bug was removed at some point in time.
It may then be unclear which version of code and data was used to produce some reported set of results.
2.	*Copy-and-paste errors*; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result.
3.	*Undocumented or ambiguous order of computation*; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another).
4.	*Ambiguous software dependencies*; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.

@@

*Version control* prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the lifetime (and in particular, in the development phase) of a research project.
Version control allows a clear link between which results were generated by which version of code and data.
This is important for reproducibility because only if it is obvious which version of data and code produced which results, those results can be said to be reproducible.
We recommend using *Git* for version control, because of its widespread adoption in the R community.

@@

Dynamic document generation integrates both steps.
Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allowing interested readers to gain a deeper understanding of the contents [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
*R Markdown* uses `Markdown` for formatting text and R (or other programming languages) for writing the statistical analysis.
`Markdown` is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
This way, `Markdown` does not need any specialized software for editing.
It is highly fool-proof [unlike, for example, `LaTeX` @lamportLATEXDocumentPreparation1994], works well with version control systems and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
We suggest using `Markdown` for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
`R Markdown` extends regular `Markdown` by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a `Markdown` document.
Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
This allows the creation of (conditionally) formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
@xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of `R Markdown`.

One would hope that sharing a dynamic document would allow individuals (including one's future self) to download and reproduce a given analysis simply.
While version control and dynamic document generation are becoming more common, we have argued that two more components are required and that each component alone is unlikely to guarantee reproducibility [@Peikert2019; @vanlissa2020worcs].
In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.

To automatically resolve dependencies between project files, we rely on *dependency tracking*.
In essence, researchers provide a collection of *computational recipes*.
A computational recipe describes how inputs are processed to deterministically create a specific output in a way that is automatically executable.
The concept of computational recipes is central to our understanding of reproducibility because it enables a unified way to reproduce a project automatically.
Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).
One noteworthy property of computational recipes is that they require a textual representation and the ability of a program to make sense of this textual representation.
A workflow that includes, e.g. renaming variable names by hand in a graphical spreadsheet application is therefore incompatible with automated reproducibility.
Another property of a computational recipe is that the same inputs should always result in the same outputs.
For most computer code (given the same software is used), this property is fulfilled.
However, one noteworthy exception is pseudo-random numbers.
Whenever random numbers are used in a computation, this computation is only reproducible when the same random numbers are used.
To ensure identical random numbers, users may fix the state of the random number generated with a so-called seed (e.g. in R with `set.seed()`), but they also need to guarantee that the pseudo-random number generator is unchanged [see @Peikert2019].

@@

There are two targets, the final rendered report (`manuscript.pdf`, l.1) and the simulation results (`simulation_results.csv`, l.4).
Each target is followed by a colon and a list of requirements.
If a requirement is newer than the target, the recipe will be executed to rebuild the target.
If a requirement does not exist, `Make` uses a recipe to build the requirement first before building the target.
Here, if one were to build the final `manuscript.pdf` by rendering the `R Markdown` with the command shown in l.2, `Make` would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation first before rendering the manuscript.
This ensures that the simulated data are present before the manuscript is built and that the simulation is re-run, and the manuscript is rebuilt if the simulation code was changed.
`Make` therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] does not explicitly contain this dependency tracking element, but its strict structure of only containing one definite `R Markdown` still makes dependencies between files unambiguous.

@@

A version-controlled dynamic document with dependency tracking may still rely on external software.
Troubleshooting issues specific to a particular programming language or dependent tool typically require considerable expertise and threatens reproducibility.
One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide all dependent software packages and system libraries via *software management*.
One comprehensive approach to software management is containerization.
The central idea is that by
"packaging the key elements of the computational environment needed to run the desired software the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p. 174].

@@

Note that the software environment of the Docker image is completely separate from the software installed on your computer.
This separation is excellent for reproducibility but takes some getting used to.
For example, it is important to realize that software available on your local computer will *not* be accessible within the confines of the `Docker` image.
Each dependency that you want to use within the `Docker` image must be explicitly added as a dependency.
Furthermore, using `Docker` may require you to install software on an operating system that may not be familiar to you.
The images supplied by @boettigerIntroductionRockerDocker2017, for example, are based on Linux.

There are two ways to build a `Docker` image.
First, users can manually install whatever software they like from within the virtual environment.
Such a manually build environment can still be ported to all computers which support `Docker`.
We, however, prefer the second way of building images automatically from a textual description called `Dockerfile`.
Because the `Dockerfile` clearly describes which software is installed how, the installation process can be repeated automatically.
Users can, therefore, quickly change the software environment, e.g. update to another R version or given package version.

Packaging all required software in such an image requires considerable amounts of storage space.
Two major strategies help to keep the storage requirements reasonable.
One strategy is to rely on pre-made images, which are maintained by a community for particular purposes.
For example, there are pre-made images that only include what is necessary for R, based on Ubuntu containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
The image for this project includes Ubuntu, R, RStudio, LaTeX as well as a variety of R packages like tidyverse [@tidyverse] and all its dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
A second strategy is to save a container recipe, a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment instead of the software environment itself.
`Dockerfiles` are tiny (the [`Dockerfile`](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
For proper archival, we, therefore, recommend storing a complete image of the software environment, in addition to the `Dockerfile`.
A more comprehensive overview of the use of containerization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] relies for software management on the R package `renv`[@R-renv], which is more lightweight and easier to use than `Docker` on the one hand, but not as comprehensive on the other hand because it only snapshots the R packages instead of all software used.

To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using `R Markdown`), version control (using `Git`), dependency tracking (using `Make`), and Software management (using `Docker`).
While `R Markdown` and `Git` are well integrated into the R environment through RStudio, `Make` and `Docker` require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable obstacle to the acceptance and implementation of the workflow.
To remove this obstacle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
These projects will remain reproducible irrespective of whether `repro` remains accessible in future.
Users do not have to have `repro` installed to reproduce a project;
in fact, they do not even need to have R installed because the entire project can be rebuilt inside a container with R installed.
In the remainder, we will walk you through creating a reproducible research project with the package `repro`.
